\documentclass[twoside]{article}
\setlength{\oddsidemargin}{0.2 in}
\setlength{\evensidemargin}{0.2in}
\setlength{\topmargin}{-0.8 in}
\setlength{\textwidth}{6.5 in}
\setlength{\textheight}{8.5 in}
\setlength{\headsep}{0.75 in}
\setlength{\parindent}{0 in}
\setlength{\parskip}{0.1 in}

%\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[tbtags]{amsmath}
\newtheorem{thm}{Theorem}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{hyperref}
\usepackage{setspace}

\begin{document}
\begin{center}
\begin{spacing}{1.75}
{\Large { ECE368: Probabilistic Reasoning} \\
	{Lab 2: Classification with Gaussian Models and Bayesian Linear Regression}}
\end{spacing}
\vspace{-1em}
\end{center}
%\vspace{-11mm}
You can complete this lab in a group of two. Please provide the name and student number of both members.
\begin{center}
   	\framebox{%
		\vbox{%
			\vskip 3mm
			\hbox to 6.28in{\rlap{\bf Name: Bharat Bhargava}\hfil{\bf Student Number: 1010380892}\hfil}
			\vskip 5mm
			\hbox to 6.28in{\rlap{\bf Name: Daivik Dhar}\hfil{\bf Student Number: 1010214260}\hfil}
			\vskip 3mm
		}%
	}
\end{center}
   
{\bf You should hand in:} 1) A scanned \textsf{.pdf} version of this sheet with your answers (file size
should be under $2$ MB); 2) Two figures for Part 1 Question 1 (c), four figures for Part 2 Question 2 and three figures for Part 2 Question 4
in the \textsf{.pdf} format; and 3) Two Python files \textsf{ldaqda.py} and \textsf{regression.py} that
contain your code. All these files should be uploaded to Quercus.

\section{Classification with Gaussian Models}
\begin{enumerate}
\item
\begin{enumerate}
\item Write down the maximum likelihood estimates of the parameters
$\boldsymbol \mu_{1}$, $\boldsymbol \mu_{2}$, $\boldsymbol \Sigma$,
$\boldsymbol \Sigma_1$, and $\boldsymbol \Sigma_2$ as functions of the training
data $\{\mathbf{x}_n, y_n\}, n=1,2,\ldots,N$. ({\bf $1$ pt})
\begin{center}
   \framebox{
    \begin{minipage}{5.5in}
   \vspace{4mm}
   
   \textbf{Class Means:} The ML estimates are simply the sample averages of the feature vectors for each respective class, filtered using the indicator function $I(\cdot)$.
   \begin{align*}
   \hat{\mu}_1 &= \frac{\sum_{n=1}^N I(y_n = 1) x_n}{\sum_{n=1}^N I(y_n = 1)}, \quad \hat{\mu}_2 = \frac{\sum_{n=1}^N I(y_n = 2) x_n}{\sum_{n=1}^N I(y_n = 2)}
   \end{align*}
   
   \textbf{QDA Covariances:} For QDA, each class has its own distinct covariance matrix, calculated using only the squared deviations from that specific class.
   \begin{align*}
   \hat{\Sigma}_1 &= \frac{\sum_{n=1}^N I(y_n = 1) (x_n - \hat{\mu}_1)(x_n - \hat{\mu}_1)^T}{\sum_{n=1}^N I(y_n = 1)} \\[6pt]
   \hat{\Sigma}_2 &= \frac{\sum_{n=1}^N I(y_n = 2) (x_n - \hat{\mu}_2)(x_n - \hat{\mu}_2)^T}{\sum_{n=1}^N I(y_n = 2)}
   \end{align*}
   
   \textbf{LDA Shared Covariance:} For LDA, both classes share the exact same covariance matrix. This is calculated as the pooled sample covariance averaged across all $N$ samples.
   \begin{align*}
   \hat{\Sigma} &= \frac{1}{N} \sum_{n=1}^N \Big[ I(y_n = 1) (x_n - \hat{\mu}_1)(x_n - \hat{\mu}_1)^T \\
   &\qquad\qquad\qquad + I(y_n = 2) (x_n - \hat{\mu}_2)(x_n - \hat{\mu}_2)^T \Big]
   \end{align*}
   
   \vspace{2mm}
   \end{minipage}
   }
   \end{center}
   \newpage
 \vspace{2mm}
\item In the case of LDA, write down the decision boundary as a linear equation of
$\mathbf{x}$ with parameters $\boldsymbol \mu_{1}$, $\boldsymbol \mu_{2}$, and $\boldsymbol \Sigma$. Note that we assume $\pi=0.5$. ({\bf $0.5$ pt})
\begin{center}
   \framebox{
      \vbox{\vspace{25mm}
    \hbox to 5.5in {}
      \vspace{20mm}}
   }
   \end{center}
In the case of QDA, write down the decision boundary as a quadratic equation of
$\mathbf{x}$ with parameters $\boldsymbol \mu_{1}$, $\boldsymbol \mu_{2}$, $\boldsymbol \Sigma_1$, and $\boldsymbol \Sigma_2$. Note that we assume $\pi=0.5$. ({\bf $0.5$ pt})
\begin{center}
   \framebox{
      \vbox{\vspace{25mm}
    \hbox to 5.5in {}
      \vspace{20mm}}
   }
   \end{center}

\item Complete function \textsf{discrimAnalysis} in \textsf{ldaqda.py} to
	visualize LDA and QDA models and the corresponding decision boundaries. Please name the figures as \textsf{lda.pdf}, and \textsf{qda.pdf}. ({\bf $1$ pt})

\end{enumerate}
\item The misclassification rates are \framebox{\vbox{\vspace{3mm}\hbox to 16mm {}}} for LDA, and \framebox{\vbox{\vspace{3mm}\hbox to 16mm {}}} for QDA. ({\bf $1$ pt})
\end{enumerate}


\section{Bayesian Linear Regression}
\begin{enumerate}
\item Express posterior distribution $p(\mathbf{a}|z_1,\ldots,z_{N} ; x_1, \ldots, x_N)$ using $\sigma^2, \beta$, $x_1, z_1, x_2, z_2, \ldots, x_N, z_N$.
   ({\bf $1$ pt})

      \begin{center}
   \framebox{
      \vbox{\vspace{25mm}
    \hbox to 5.5in {}
      \vspace{10mm}}
   }
   \end{center}
\item Let $\sigma^2=0.1$ and $\beta=1$. Draw four contour plots corresponding to the distributions $p(\mathbf{a})$, $p(\mathbf{a}|z_1;x_1)$, $p(\mathbf{a}|z_1,\ldots, z_{5}; x_1 \ldots x_5)$, and $p(\mathbf{a}| z_1,\ldots,z_{100}; x_1 \ldots x_{100})$. In all contour plots, the x-axis represents $a_0$, and the y-axis represents $a_1$. Please save the figures with names \textbf{prior.pdf, posterior1.pdf, posterior5.pdf, posterior100.pdf}, respectively. ({\bf $1.5$ pt})

\item Suppose that there is a new input $x$, for which we want to predict the corresponding target value $z$. Write down the distribution of the prediction $z$, i.e, $p(z|z_1,\ldots,z_N ; x, x_1, \ldots x_N)$. ({\bf $1$ pt})
     \begin{center}
   \framebox{
      \vbox{\vspace{30mm}
    \hbox to 5.5in {}
      \vspace{5mm}}
   }
   \end{center}


\item Let $\sigma^2=0.1$ and $\beta=1$. Given a set of new inputs $\{-4,-3.8,\ldots,3.8,4\}$, plot three figures, whose x-axis is the input and y-axis is the prediction, corresponding to three cases:
    \begin{enumerate}
    \item The predictions are based on one training sample, i.e., based on $p(z|z_1;x,x_1,)$.
    \item The predictions are based on $5$ training samples, i.e., based on $p(z|z_1,\ldots,z_{5};x,x_1,\ldots,x_{5})$.
    \item The predictions are based on $100$ training samples, i.e., based on  $p(z|z_1,\ldots,z_{100};x,x_1,\ldots,x_{100})$.
    \end{enumerate}

     The range of each figure is set as $[-4, 4]\times [-4, 4]$. Each figure should contain the following three components: 1) the new inputs and the corresponding predicted targets; 2) a vertical interval at each predicted target, indicating the range within one standard deviation; 3) the training sample(s) that are used for the prediction. Use \textsf{plt.errorbar} for 1) and 2); use \textsf{plt.scatter} for 3). Please save the figures with names \textbf{predict1.pdf, predict5.pdf, predict100.pdf}, respectively. ({\bf $1.5$ pt})

\end{enumerate}

\end{document}
